A comparison of classification algorithms applied to the Adult Database

### 2009-12-16 demo/demo.tex
c4.5, random forest and bayes slides

  \begin{itemize}
    \item<1-> Classifiers were built using the train set and then tested against the test set
    \item<1-> C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
    \begin{itemize}
      \item<1-> Various parameters available in Weka algorithms were experimented in hope of getting better error rates
      \item<1-> Every missing value was replaced with modes and means
      \item<1-> Two different classifiers were trained for each algorithm, one using only discretized attributes
                and unchanged attributes to analyze how the algorithms behaves when using numeric attributes
    \end{itemize}

@@

@@

    \begin{itemize}
    \item<1-> ID3 and C4.5 are very fast and build classifiers with a decent error rate.
    \item<1-> ILP and CN2 build more expressive classifiers, capable of a better interpretation by a human agent, but take longer to finish.
    \item<1-> The worst error rate is below 25\%.
    \item<1-> There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.
    \begin{itemize}
        \item<1-> Open source.
        \item<1-> Easy to use.
        \item<1-> APIs for integration with Java applications.
    \end{itemize}
    \end{itemize}
}

>>>>

  \begin{itemize}
    \item<1-> Classifiers were built using the train set and then tested against the test set
    \item<1-> C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
    \begin{itemize}
      \item<1-> Various parameters available in Weka algorithms were experimented in hope of getting better error rates
      \item<1-> Every missing value was replaced with modes and means
      \item<1-> Two different classifiers were trained for each algorithm, one using only discretized attributes
                and another with the attributes unmodified
      \item<1-> Analyze how the algorithm behaves when using numeric attributes
    \end{itemize}
  \end{itemize}
}

@@

\subsection{C4.5}
\frame
{
  \frametitle{C4.5}

  \begin{itemize}
    \item Weka supports a confidence threshold parameter that dictates how much pruning the algorithm will do
    \item The bigger the threshold the less pruning the algorithm does
    \item When using only discretized data the algorithm gives error rates of $\approx$25\%
  \end{itemize}

  \begin{table}[ht]
    \begin{center}
    \scalebox{0.8}{
    \begin{tabular}{ | l | c | c | c | c | c |}
      \hline
      \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
      -U & 9.48 & 21.58 & 0m7.525s & 6422 & 5491 \\ \hline
      -C 0.05 & 13.78 & 19.49 & 0m7.667s & 149 & 105 \\ \hline
      -C 0.1 & 13.37 & 19.06 & 0m7.969s & 278 & 212 \\ \hline
      -C 0.15 & 12.99 & 18.84 & 0m7.708s & 361 & 278 \\ \hline
      -C 0.2 & 12.95 & 18.89 & 0m7.812s & 373 & 284 \\ \hline
      -C 0.25 & 12.86 & 18.86 & 0m7.974s & 433 & 331 \\ \hline
      -C 0.3 & 12.11 & 19.73 & 0m8.064s & 806 & 624 \\ \hline
      -C 0.35 & 12.02 & 19.77 & 0m8.035s & 878 & 678 \\ \hline
      -C 0.4 & 11.54 & 20.47 & 0m8.255s & 1264 & 995 \\ \hline
      -C 0.5 & 11.22 & 20.28 & 0m8.117s & 1603 & 1257 \\ \hline
    \end{tabular}}
    \end{center}
  \end{table}
}

\subsection{Random Forest}
\frame
{
  \frametitle{Random Forest}

  \begin{itemize}
    \item<1-> It is an ensemble classifier consisting of many decision trees
    \item<1-> Outputs the class that is the mode of the class's output by individual trees
    \item<1-> Better results require more time building the classifier
    \item<1-> Large quantities of memory needed: 1GB to compute with 400 trees
    \item<1-> Over-fitting
  \end{itemize}

  \begin{table}[ht]
    \begin{center}
    \scalebox{0.75}{
    \begin{tabular}{ | l | c | c | c |}
      \hline
      \textbf{Trees} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
      2 & 5.72 & 25.70 & 0m4.882s \\ \hline
      4 & 3.81 & 23.40 & 0m6.086s \\ \hline
      5 & 3.20 & 23.43 & 0m6.839s \\ \hline
      7 & 2.86 & 21.01 & 0m8.190s \\ \hline
      10 & 2.69 & 21.15 & 0m10.162s \\ \hline
      20 & 2.46 & 20.83 & 0m16.970s \\ \hline
      40 & 2.44 & 21.26 & 0m30.488s \\ \hline
      80 & 2.44 & 21.70 & 1m1.347s \\ \hline
      100 & 2.44 & 21.75 & 1m15.592s \\ \hline
      400 & 2.44 & 21.62 & 7m51.624s \\ \hline
    \end{tabular}}
    \end{center}
  \end{table}
}

\subsection{Bayesian Networks}
\frame
{
  \frametitle{Bayesian Networks}

  \begin{itemize}
    \item<1-> Based on conditional probabilities and the Bayes theorem
    \item<1-> Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables
    \item<1-> The algorithm first learns the network structure / graph and then learns the probability tables
    \item<1-> We experimented with different algorithms to learn the network structure:
     \begin{itemize}
      \item<1-> Genetic Search
      \item<1-> Hill Climbing
      \item<1-> Simulated Annealing
     \end{itemize}
    \item<1-> Weka provides a very rich set of parameters for each search algorithm
  \end{itemize}

}

@@

    \begin{itemize}
    \item<1-> ID3 and C4.5 are very fast and build classifiers with a decent error rate.
    \item<1-> ILP and CN2 build more expressive classifiers, capable of a better interpretation by a human agent, but take longer to finish.
    \item<1-> The worst error rate is below 25\%.
    \item<1-> Using numeric information results in better error rates
    \item<1-> There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.
    \begin{itemize}
        \item<1-> Open source.
        \item<1-> Easy to use.
        \item<1-> APIs for integration with Java applications.
    \end{itemize}
    \end{itemize}
}
